### 2019-09-27 vesicle_code/docs/softmatter/permeableVes.tex
made changes to soft matter Tex file.  Area eqs and Num sims section

We can use the governing equations to write a differential equation for
the area of the vesicle as a function of time. The area inside a single
vesicle $\gamma$ parameterized with $\mbx$ is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\mbx \cdot \mbn)\, ds = 
    \frac{1}{2}\int_{0}^{2\pi} \left(\mbx \cdot
    \pderiv{\mbx}{\theta}^\perp\right)
    \, d\theta.
\end{align}
Taking the time derivative of $A(t)$,
\begin{align}
  \dot{A}(t) &= \frac{1}{2}\int_{0}^{2\pi} \left(\dot{\mbx} \cdot
  \pderiv{\mbx}{\theta}^{\perp} + 
    \mbx \cdot \pderiv{\dot{\mbx}}{\theta}^{\perp}\right) d\theta, \\
  &= \frac{1}{2}
  \int_{0}^{2\pi}(\dot{\mbx}\cdot\mbn)\left\|\pderiv{\mbx}{\theta}\right\|d\theta +
     \frac{1}{2}\int_{0}^{2\pi}
    \left(\mbx \cdot \pderiv{\dot{\mbx}}{\theta}^{\perp}\right)d\theta, \\
  &= \frac{k_w}{2} \int_{\gamma}\left(\mbf\cdot\mbn\right)ds +
  \frac{1}{2}\int_{0}^{2\pi}\left(\mbx \cdot \pderiv{}{t}\left(\mbn
  \left\|\pderiv{\mbx}{\theta}\right\|\right)\right)d\theta, \\
  &= \frac{k_w}{2} \int_{\gamma}\left(\mbf\cdot\mbn\right)ds + \frac{1}{2}
  \int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right)ds + \frac{1}{2}\int_{0}^{2\pi}
  \frac{\mbx\cdot\mbn}{\left\|\pderiv{\mbx}{\theta}\right\|}
  \left(\pderiv{\mbx}{\theta}\cdot
  \pderiv{\dot{\mbx}}{\theta}\right)d\theta \\
  &= \frac{k_w}{2} \int_{\gamma}\left(\mbf\cdot\mbn\right)ds + \frac{1}{2}
  \int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right)ds + \frac{1}{2}\int_{\gamma}
  (\mbx\cdot\mbn) (\mbx_s \cdot \dot{\mbx}_s)\, ds.
\end{align}
To satisfy the inextensibilty condition, $\mbx_s \cdot \dot{\mbx}_s = 0$.
Therefore the time derivative of $A(t)$ is
\begin{align}
     \dot{A}(t) &= \frac{k_w}{2}
     \int_{\gamma}\left(\mbf\cdot\mbn\right)ds
     + \frac{1}{2} \int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right)ds.
\end{align}
... 
\begin{align}
    \dot{A}(t) &= \frac{k_w}{2}
    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds+\frac{1}{2}\int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right) ds,
\end{align}
which we re-write as
\begin{align}
    \dot{A}(t) &= \frac{k_w}{2}
    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds-\frac{1}{2}\int_{\gamma}\left(\mbx^{\perp}\cdot\dot{\mbx}_s\right) ds.
\end{align}
Applying integration by parts,
\begin{align}
    \dot{A}(t) &= \frac{k_w}{2}
    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds+\frac{1}{2}\int_{\gamma}\left(\mbx^{\perp}_s\cdot\dot{\mbx}\right) ds,
\end{align}
which we re-write as
\begin{align}
    \dot{A}(t) &= \frac{k_w}{2}
    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds+\frac{1}{2}\int_{\gamma}\left(\dot{\mbx}\cdot\mbn\right),\\
    \dot{A}(t) &= k_w \int_{\gamma}\left(\mbf\cdot\mbn\right)ds.
\end{align}
At steady stae,
\begin{align}
    \dot{A}(t) &= k_w \int_{\gamma}\left(\mbf\cdot\mbn\right)ds =0.
\end{align}
We now replace $\mbf$ with its relation
\begin{align}
   k_w\int_{\gamma}
   \left((-\kappa_b\mbx_{ssss}+(\sigma\mbx_s)_s)\cdot\mbn\right)ds &=0.
\end{align}
Applying integration by parts,
\begin{align}
     k_w\int_{\gamma}\left(\kappa_b\mbx_{sss}-\sigma\mbx_s\right)\cdot
     \mbn_s ds &=0.
\end{align}
From the Frenet-Serret formulas, we have 
\begin{align}
     -k_w\int_{\gamma}
     \left(\kappa_b\mbx_{sss}-\sigma\mbx_s\right)\cdot\kappa\mbx_s ds &=0.
\end{align}

Applying integration by parts and Frenet-Serret again
\begin{align}
     k_w\int_{\gamma}
     \kappa_b\kappa\left(\mbx_{ss}\cdot\mbx_{ss}\right)ds + k_w
     \int_{\gamma} \kappa_b \kappa_s(\mbx_{ss} \cdot \mbx_s)ds +k_w\int_{\gamma}\kappa\sigma ds  &= 0,\\
    k_w\int_{\gamma} \kappa_b\kappa(\kappa\mbn \cdot
     \kappa\mbn)+ k_w \int_{\gamma} \kappa_b \kappa_s(\mbx_{ss} \cdot
     \mbx_s)ds +k_w\int_{\gamma}\kappa\sigma ds &= 0,
\end{align}
which we re-write as
\begin{align}
   k_w\int_{\gamma}\kappa_b\kappa^3 ds + k_w \int_{\gamma} \kappa_b
   \kappa_s(\mbx_{ss} \cdot \mbx_s)ds +k_w\int_{\gamma} \kappa\sigma ds = 0.
\end{align}
Applying Frenet-serret again,
\begin{align}
    k_w\int_{\gamma}\kappa_b\kappa^3 ds + k_w \int_{\gamma} \kappa_b
    \kappa_s(\kappa\mbn \cdot \TT)ds +k_w\int_{\gamma} \kappa\sigma ds = 0, \\
    k_w\int_{\gamma}\kappa_b\kappa^3 ds +k_w\int_{\gamma} \kappa\sigma ds  = 0, 
\end{align}
we compare this to our numerical results.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
To address stiffness associated with high-order derivatives,
equation~\eqref{eqn:vesVelocity} is discretized with a semi-implicit
time stepping method similar to our previous works.  First, we introduce
the notation
\begin{align}
  B[\mbx]\mbf &= -\kappa_b\frac{d^4}{ds^4} \mbf, \\
  T[\mbx]\sigma &= (\sigma \mbx_s)_s, \\
  P[\mbx]\mbf &= (\mbf \cdot \mbn) \mbn, \\
  D[\mbx]\mbf &= \mbx_s \cdot \mbf_s,
\end{align}
where $\mbn$ is the outward unit normal of the vesicle
parameterized by $\mbx$, and $d/ds$ is its arclength derivative.
Dropping the dependence of the operators on $\mbx$, the dynamics of the
vesicle are governed by
\begin{align}
  \dot{\mbx} &= \mbu_{\infty}(\mbx) + 
  \beta P(B\mbx + T\sigma) + \sgl(B\mbx + T\sigma), \\
  D \dot{\mbx} &= 0.
\end{align}
Using the notation $B^N$ to denote the bending operator due to the
vesicle configuration at time $t_N$, and similar notation for the other
operators, we apply the semi-implicit time stepping method
\begin{align}  
  \frac{\mbx^{N+1} - \mbx^N}{\Delta t} = \mbu_\infty(\mbx^N) 
  + \beta P^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}) 
  + \sgl^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}).
\end{align}
Applying $D$ to the first-order time discretization of
equation~\eqref{eqn:vesVelocity}, the inextensibility condition is
\begin{align*}
  &D^N(\mbx^{N+1} - \mbx^N) = 0 \\ 
  \Rightarrow &D^N\mbx^{N+1} = D^N \mbx^N \\ 
  \Rightarrow &D^N\mbx^{N+1} = 1.
\end{align*}
Finally, we write the time stepping method in matrix form as
\begin{align}
  A \left(
    \begin{array}{c}
      \mbx^{N+1} \\ \sigma^{N+1}
    \end{array}
  \right) = 
  \left(
    \begin{array}{c}
      \mbx^{N} + \Delta t \mbu_\infty(\mbx^N) \\ \mathbf{1}
    \end{array}
  \right),
\end{align}
where
\begin{align}
 A = \left(
  \begin{array}{cc}
    I + \Delta t \beta P^N B^N + \Delta t \sgl^N B^N & 
    -\Delta t P^N T^N - \Delta t \sgl^N T^N \\
    D^N & 0
  \end{array}
  \right).
\end{align}

@@

  \item For $\beta$ that depends on arclength, we have observed both a
    breathing motion (semi-permeable only on opposite sides), and
    migration (semi-permeable only on about 25\% of the vesicle).
\end{itemize}

>>>>

We can use the governing equations to write a differential equation for
the area of the vesicle as a function of time. The area inside a single
vesicle $\gamma$ parameterized with $\mbx$ is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\mbx \cdot \mbn)\, ds = 
    \frac{1}{2}\int_{0}^{2\pi} \left(\mbx \cdot
    \pderiv{\mbx}{\theta}^\perp\right)
    \, d\theta.
\end{align}
Taking the time derivative of $A(t)$,
\begin{align}
  \dot{A}(t) &= \frac{1}{2}\int_{0}^{2\pi} \left(\dot{\mbx} \cdot
  \pderiv{\mbx}{\theta}^{\perp} + 
    \mbx \cdot \pderiv{\dot{\mbx}}{\theta}^{\perp}\right) d\theta, \\
  %&= \frac{1}{2}
  %\int_{0}^{2\pi}(\dot{\mbx}\cdot\mbn)\left\|\pderiv{\mbx}{\theta}\right\|d\theta +
  %   \frac{1}{2}\int_{0}^{2\pi}
  %  \left(\mbx \cdot \pderiv{\dot{\mbx}}{\theta}^{\perp}\right)d\theta, \\
  %&= \frac{k_w}{2} \int_{\gamma}\left(\mbf\cdot\mbn\right)ds +
  %\frac{1}{2}\int_{0}^{2\pi}\left(\mbx \cdot \pderiv{}{t}\left(\mbn
  %\left\|\pderiv{\mbx}{\theta}\right\|\right)\right)d\theta, \\
  %&= \frac{k_w}{2} \int_{\gamma}\left(\mbf\cdot\mbn\right)ds + \frac{1}{2}
  %\int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right)ds + \frac{1}{2}\int_{0}^{2\pi}
  %\frac{\mbx\cdot\mbn}{\left\|\pderiv{\mbx}{\theta}\right\|}
  %\left(\pderiv{\mbx}{\theta}\cdot
  %\pderiv{\dot{\mbx}}{\theta}\right)d\theta \\
  &= \frac{k_w}{2} \int_{\gamma}\left(\mbf\cdot\mbn\right)ds + \frac{1}{2}
  \int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right)ds + \frac{1}{2}\int_{\gamma}
  (\mbx\cdot\mbn) (\mbx_s \cdot \dot{\mbx}_s)\, ds.
\end{align}
To satisfy the inextensibilty condition, $\mbx_s \cdot \dot{\mbx}_s = 0$, the time derivative of $A(t)$ is
\begin{align}
     \dot{A}(t) &= \frac{k_w}{2}
     \int_{\gamma}\left(\mbf\cdot\mbn\right)ds
     + \frac{1}{2} \int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right)ds.
\end{align}
%... 
%\begin{align}
%    \dot{A}(t) &= \frac{k_w}{2}
%    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds+\frac{1}{2}\int_{\gamma}\left(\mbx\cdot\dot{\mbn}\right) ds,
%\end{align}
which we re-write as
\begin{align}
    \dot{A}(t) &= \frac{k_w}{2}
    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds-\frac{1}{2}\int_{\gamma}\left(\mbx^{\perp}\cdot\dot{\mbx}_s\right) ds.
\end{align}
Applying integration by parts and re-writing,
%\begin{align}
%    \dot{A}(t) &= \frac{k_w}{2}
%    \int_{\gamma}\left(\mbf\cdot\mbn\right)ds+\frac{1}{2}\int_{\gamma}\left(\mbx^{\perp}_s\cdot\dot{\mbx}\right) ds,
%\end{align}
%which we re-write as
\begin{align}
 %   \dot{A}(t) &= \frac{k_w}{2}
 %   \int_{\gamma}\left(\mbf\cdot\mbn\right)ds+\frac{1}{2}\int_{\gamma}\left(\dot{\mbx}\cdot\mbn\right),\\
    \dot{A}(t) &= k_w \int_{\gamma}\left(\mbf\cdot\mbn\right)ds.
\end{align}
At steady state,
\begin{align}
    \dot{A}(t) &= k_w \int_{\gamma}\left(\mbf\cdot\mbn\right)ds =0.
\end{align}
Replacing $\mbf$ with its relation, applying integration by parts twice and using Frenet-serret formulas,  
\begin{align}
k_w\int_{\gamma}\kappa_b\kappa^3 ds +k_w\int_{\gamma} \kappa\sigma ds  = 0. 
\end{align} 
%We now replace $\mbf$ with its relation
%\begin{align}
%   k_w\int_{\gamma}
%   \left((-\kappa_b\mbx_{ssss}+(\sigma\mbx_s)_s)\cdot\mbn\right)ds &=0.
%\end{align}
%Applying integration by parts,
%\begin{align}
%     k_w\int_{\gamma}\left(\kappa_b\mbx_{sss}-\sigma\mbx_s\right)\cdot
%     \mbn_s ds &=0.
%\end{align}
%From the Frenet-Serret formulas, we have 
%\begin{align}
%     -k_w\int_{\gamma}
%     \left(\kappa_b\mbx_{sss}-\sigma\mbx_s\right)\cdot\kappa\mbx_s ds &=0.
%\end{align}

%Applying integration by parts and Frenet-Serret again
%\begin{align}
%     k_w\int_{\gamma}
%     \kappa_b\kappa\left(\mbx_{ss}\cdot\mbx_{ss}\right)ds + k_w
%     \int_{\gamma} \kappa_b \kappa_s(\mbx_{ss} \cdot \mbx_s)ds +k_w\int_{\gamma}\kappa\sigma ds  &= 0,\\
%    k_w\int_{\gamma} \kappa_b\kappa(\kappa\mbn \cdot
%     \kappa\mbn)+ k_w \int_{\gamma} \kappa_b \kappa_s(\mbx_{ss} \cdot
%     \mbx_s)ds +k_w\int_{\gamma}\kappa\sigma ds &= 0,
%\end{align}
%which we re-write as
%\begin{align}
%   k_w\int_{\gamma}\kappa_b\kappa^3 ds + k_w \int_{\gamma} \kappa_b
%   \kappa_s(\mbx_{ss} \cdot \mbx_s)ds +k_w\int_{\gamma} \kappa\sigma ds = 0.
%\end{align}
%Applying Frenet-serret again,
%\begin{align}
%    k_w\int_{\gamma}\kappa_b\kappa^3 ds + k_w \int_{\gamma} \kappa_b
%    \kappa_s(\kappa\mbn \cdot \TT)ds +k_w\int_{\gamma} \kappa\sigma ds = 0, \\
%    k_w\int_{\gamma}\kappa_b\kappa^3 ds +k_w\int_{\gamma} \kappa\sigma ds  = 0, 
%\end{align}
%we compare this to our numerical results.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
To address stiffness associated with high-order derivatives,
equation~\eqref{eqn:vesVelocity} is discretized with a semi-implicit
time stepping method similar to our previous works.  First, we introduce
the notation
\begin{align}
  B[\mbx]\mbf &= -\kappa_b\frac{d^4}{ds^4} \mbf, \\
  T[\mbx]\sigma &= (\sigma \mbx_s)_s, \\
  P[\mbx]\mbf &= (\mbf \cdot \mbn) \mbn, \\
  D[\mbx]\mbf &= \mbx_s \cdot \mbf_s,
\end{align}
where $\mbn$ is the outward unit normal of the vesicle
parameterized by $\mbx$, and $d/ds$ is its arclength derivative.
Dropping the dependence of the operators on $\mbx$, the dynamics of the
vesicle are governed by
\begin{align}
  \dot{\mbx} &= \mbu_{\infty}(\mbx) + 
  \beta P(B\mbx + T\sigma) + \sgl(B\mbx + T\sigma), \\
  D \dot{\mbx} &= 0.
\end{align}
Using the notation $B^N$ to denote the bending operator due to the
vesicle configuration at time $t_N$, and similar notation for the other
operators, we apply the semi-implicit time stepping method
\begin{align} 
\begin{split}
  \frac{\mbx^{N+1} - \mbx^N}{\Delta t} = \mbu_\infty(\mbx^N) 
  + \beta P^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}) \\
   + \sgl^N(B^N\mbx^{N+1} + T^N\sigma^{N+1}).
   \end{split} 
\end{align}
Applying $D$ to the first-order time discretization of
equation~\eqref{eqn:vesVelocity}, the inextensibility condition is
\begin{align*}
  &D^N(\mbx^{N+1} - \mbx^N) = 0 \\ 
  \Rightarrow &D^N\mbx^{N+1} = D^N \mbx^N \\ 
  \Rightarrow &D^N\mbx^{N+1} = 1.
\end{align*}
Finally, we write the time stepping method in matrix form as
\begin{align}
  A \left(
    \begin{array}{c}
      \mbx^{N+1} \\ \sigma^{N+1}
    \end{array}
  \right) = 
  \left(
    \begin{array}{c}
      \mbx^{N} + \Delta t \mbu_\infty(\mbx^N) \\ \mathbf{1}
    \end{array}
  \right),
\end{align}
where
\begin{align}
 A = \left(
  \begin{array}{cc}
    I + \Delta t \beta P^N B^N + \Delta t \sgl^N B^N & 
    -\Delta t P^N T^N - \Delta t \sgl^N T^N \\
    D^N & 0
  \end{array}
  \right).
\end{align}

@@

  \item For $\beta$ that depends on arclength, we have observed both a
    breathing motion (semi-permeable only on opposite sides), and
    migration (semi-permeable only on about 25\% of the vesicle).

  \item Compare area results to analytic equation?
\end{itemize}
